\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-10-03 gwnnDraft02.tex
adjusted table and edits

\section{Introduction} 
\label{sec:intro} 
The recent experimental detection of gravitational waves~\cite{Abbott:2016blz} gives humanity a new way to observe the universe. Future experiments~\cite{Danzmann:1994mma,Seto:2001qf,Crowder:2005nr,PhysRevD.72.083005,Harry:2006fi,Janssen:2014dka,Audley:2017drz,2017arXiv170200786A,Sato_2017,10.1093/ptep/pty078} will greatly expand the frequency range observable. Thus far, experiments have only observed recent events such as black hole mergers, but phase transitions in the early universe can leave an imprint as a \textit{stochastic} gravitational wave background~\cite{Witten:1984rs,Hogan:1984hx,Hogan:1986qda,PhysRevLett.65.3080,Caprini:2015zlo,Mazumdar:2018dfl}. Thus, searches for this background of gravitational waves can give direct information of the history of the universe before big bang nucleosynthesis. 
Because gravity is universal, gravitational waves can allow us to probe hidden sectors that couple very weakly, or not at all, to the Standard Model as long they are reheated after inflation. This was first explored in~\cite{Schwaller:2015tja}, and there has been significant work on this idea since~\cite{Jaeckel:2016jlh,Addazi:2016fbj,Hardy:2016mns,Dienes:2016vei,Tsumura:2017knk,Acharya:2017szw,Bernal:2017kxu,Aoki:2017aws,Heikinheimo:2018esa,Geller:2018mwu,Croon:2018erz,Baldes:2018emh,Bai:2018dxf,Breitbach:2018ddu,Fairbairn:2019xog,Helmboldt:2019pan,Caputo:2019wsd,Bertone:2019irm}.

@@

In this setup, there are two qualitatively different kinds of sectors:
\begin{itemize}
\item \textbf{Standard Sectors}: Those with $m_H^2 < 0$ where electroweak symmetry is broken by the vacuum expectation value (VEV) of a fundamental scalar. As in~\cite{Arkani-Hamed:2016rle}, we assume that the standard sector with the smallest absolute value of $m_H^2$ is the Standard Model.
\item \textbf{Exotic Sectors}: Those with  $m_H^2 > 0$. In this case, electroweak symmetry is preserved below the mass of the Higgs, and broken by the confinement of QCD~\cite{Susskind:1978ms}.
\end{itemize}
Cosmological observations, particularly limits on extra relativistic degrees of freedom at the time of Big Bang Nucleosynthesis and the time of the formation of the cosmic microwave background (CMB)~\cite{Aghanim:2018eyx}, require that most of the energy in the universe is in the Standard Model sector as we will quantify. Therefore, the hidden sectors cannot be in thermal equilibrium at any time, and the physics of reheating must dump energy preferentially in the Standard Model sector. This can be accomplished with primodial axion-like particle (ALP) models~\cite{MARSH20161,PhysRevLett.121.201303} and with the reheaton method~\cite{Arkani-Hamed:2016rle}. We will also explore alternative parameterizations of reheating that satisfy this condition.

@@

\begin{table*}[t]
\begin{center}
\begin{tabular}{ |p{3cm}||p{3cm}|p{3cm}|p{3cm}|p{3cm}|  }
 \hline
 \multicolumn{5}{|c|}{Parameters for multi-hidden sector scenarios} \\
 \hline
\multicolumn{5}{|c|}{Maximized signal} \\
 \hline
 Additional Sector & Higgs vev & $\Lambda^{ex}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & $N$naturalness index\\
 \hline
 1 & 24.6 EeV  & 38.6 & 87.7 & $10^{16}$ \\
 \hline
\multicolumn{5}{|c|}{Large split scenario} \\
 \hline
 Additional Sector & Higgs vev & $\Lambda^{ex}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & $N$naturalness index\\
 \hline
 1 & 246 PeV  & 10.8 & 30.3 & $10^{12}$ \\
 2 & 7.8 EeV  & 28.1 & 78.6 & $10^{15}$ \\
 \hline
 \multicolumn{5}{|c|}{Medium split scenario} \\
 \hline
 Additional Sector & Higgs vev & $\Lambda^{ex}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & $N$naturalness index\\
 \hline
 1 & 246 PeV  & 10.8 & 30.3 & $10^{12}$ \\
 2 & 778 PeV  & 14.9 & 41.6 & $10^{13}$ \\
 \hline
 \multicolumn{5}{|c|}{Five sector scenario} \\
 \hline
 Additional Sector & Higgs vev & $\Lambda^{ex}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & $N$naturalness index\\
 \hline
 1 & 246 PeV & 10.8 & 38.8 & $10^{12}$ \\
 2 & 426 PeV & 12.6 & 45.2 & $3*10^{12}$ \\
 3 & 778 PeV & 14.9 & 53.3 & $10^{13}$ \\
 4 & 1.3 EeV & 17.3 & 62.1 & $3*10^{13}$ \\
 5 & 2.5 EeV & 20.5 & 73.3 & $10^{14}$ \\
 \hline
\end{tabular}
\caption{Outline of parameters used for the various multi-hidden sector scenarios. The higgs vev is the vev for the given additional sector, $\Lambda^{ex}_{QCD}$ is the QCD phase transition in the additional sector, and $T_\gamma$ is the temperature of the SM photon bath when the SFOPT occurs in the additional sector.}\label{tab:cases}
\end{center}
\end{table*}

>>>>

\section{Introduction} 
\label{sec:intro} 
The recent experimental detection of gravitational waves~\cite{Abbott:2016blz} gives humanity a new way to observe the universe. Future experiments~\cite{Danzmann:1994mma,Seto:2001qf,Crowder:2005nr,PhysRevD.72.083005,Harry:2006fi,Janssen:2014dka,Audley:2017drz,2017arXiv170200786A,Sato_2017,10.1093/ptep/pty078} will greatly expand the frequency range observable. Thus far, experiments have only observed recent events such as black hole mergers, but phase transitions in the early universe can leave an imprint as a \textit{stochastic} gravitational wave background~\cite{Witten:1984rs,Hogan:1984hx,Hogan:1986qda,PhysRevLett.65.3080,Caprini:2015zlo,Mazumdar:2018dfl}. Thus, searches for this background of gravitational waves can give direct information about the history of the universe before big bang nucleosynthesis. Because gravity is universal, gravitational waves can allow us to probe hidden sectors that couple very weakly, or not at all, to the Standard Model as long they are reheated after inflation. This was first explored in~\cite{Schwaller:2015tja}, and there has been significant work on this idea since~\cite{Jaeckel:2016jlh,Addazi:2016fbj,Hardy:2016mns,Dienes:2016vei,Tsumura:2017knk,Acharya:2017szw,Bernal:2017kxu,Aoki:2017aws,Heikinheimo:2018esa,Geller:2018mwu,Croon:2018erz,Baldes:2018emh,Bai:2018dxf,Breitbach:2018ddu,Fairbairn:2019xog,Helmboldt:2019pan,Caputo:2019wsd,Bertone:2019irm}.

@@

In this setup, there are two qualitatively different kinds of sectors:
\begin{itemize}
\item \textbf{Standard Sectors}: Those with $m_H^2 < 0$ where electroweak symmetry is broken by the vacuum expectation value (vev) of a fundamental scalar. As in~\cite{Arkani-Hamed:2016rle}, we assume that the standard sector with the smallest absolute value of $m_H^2$ is the Standard Model.
\item \textbf{Exotic Sectors}: Those with  $m_H^2 > 0$. In this case, electroweak symmetry is preserved below the mass of the Higgs, and broken by the confinement of QCD~\cite{Susskind:1978ms}.
\end{itemize}
Cosmological observations, particularly limits on extra relativistic degrees of freedom at the time of Big Bang Nucleosynthesis and the time of the formation of the cosmic microwave background (CMB)~\cite{Aghanim:2018eyx}, require that most of the energy in the universe is in the Standard Model sector as we will quantify. Therefore, the hidden sectors cannot be in thermal equilibrium at any time, and the physics of reheating must dump energy preferentially in the Standard Model sector. This can be accomplished with primodial axion-like particle (ALP) models~\cite{MARSH20161,PhysRevLett.121.201303} and with the reheaton method~\cite{Arkani-Hamed:2016rle}. We will also explore alternative parameterizations of reheating that satisfy this condition.

@@

\begin{table*}[t]
\begin{center}
\begin{tabular}{ |p{2cm}||p{3cm}|p{3cm}|p{3cm}|p{2cm}|  }
 \hline
 \multicolumn{5}{|c|}{Parameters for multi-hidden sector benchmarks} \\
 \hline
\multicolumn{5}{|c|}{Maximized signal} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $24.6 \times 10^{9}$ & 38.6 & 87.7 & $10^{16}$ \\
 \hline
\multicolumn{5}{|c|}{Large split} \\
 \hline
 Sector & Higgs vev (GeV & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $246 \times 10^6$  & 10.8 & 30.3 & $10^{12}$ \\
 2 & $7.8 \times 10^9$  & 28.1 & 78.6 & $10^{15}$ \\
 \hline
 \multicolumn{5}{|c|}{Medium split} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $246 \times 10^{6}$ & 10.8 & 30.3 & $10^{12}$ \\
 2 & $778 \times 10^6$  & 14.9 & 41.6 & $10^{13}$ \\
 \hline
 \multicolumn{5}{|c|}{Five sectors} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $246 \times 10^6$ & 10.8 & 38.8 & $10^{12}$ \\
 2 & $426 \times 10^6$ & 12.6 & 45.2 & $3*10^{12}$ \\
 3 & $778 \times 10^6$ & 14.9 & 53.3 & $10^{13}$ \\
 4 & $1.3 \times 10^9$ & 17.3 & 62.1 & $3*10^{13}$ \\
 5 & $2.5 \times 10^9$ & 20.5 & 73.3 & $10^{14}$ \\
 \hline
\end{tabular}
\caption{Outline of parameters used for the various multi-hidden sector scenarios. The higgs vev is the vev for the given additional sector, $\Lambda^{AS}_{QCD}$ is the QCD phase transition in the additional sector, and $T_\gamma$ is the temperature of the SM photon bath when the SFOPT occurs in the additional sector. The index indicates the equivalent sector from the $N$naturalness model (Eq.~(\ref{eqn:vevs})).}\label{tab:cases}
\end{center}
\end{table*}
